Manuscript_Github_Pchabaudi_pir_TPMHeatmaps

### 2021-04-20 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir_TPMHeatmaps.Rmd
Heatmap separated by sub-family

```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

>>>>

```{r highest_expressed_stage_heatmap_fig3b}

cir_housekeeping_id <- c(cir_id, 
                         'PCHAS_0721000',
                         'PCHAS_0831300',
                         'PCHAS_0931000')
```

```{r}
# Get order of pir genes by highest expressed genes in each sample.

combined_data_pir <- twentyfour_avgtpm$RMT$tpm %>% 
  filter(Geneid %in% cir_housekeeping_id) %>% 
  column_to_rownames(var = 'Geneid') %>% 
  filter_all(
      all_vars(. > 1)
  )

#Ordering that a more understandable order

combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
                                                   order_vector)),
                                       drop = FALSE]

combined_data_ordered <- tibble(Geneid = names(apply(combined_data_pir, 1, mean)),
                                means = apply(combined_data_pir, 1, mean),
                                subfam = factor(
                                  ifelse(rownames(combined_data_pir) %in% cir_id,
                                         cir_info$subfam[match(rownames(combined_data_pir), 
                                                               cir_info$Geneid)],
                                         'ref')
                                )) %>% 
  mutate(subfam =
           forcats::fct_shift(
             relevel(
               as.factor(subfam),
               ref = 'ref')
           )
  ) %>% 
  # filter(means > 1) %>% 
  arrange(subfam, 
          desc(means)) %>% 
  .$Geneid

#Order the heatmap data by the order defined above (Subfam, then mean expression).
combined_data_pir_ht <- combined_data_pir[
  order(
    match(rownames(combined_data_pir),
          combined_data_ordered),
    na.last = NA #Remove NAs
  ),
]

#Change the colnames to the times
colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
                                              pattern = '[[:digit:]]{2}h')

#Order by time point
combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
                                             drop = FALSE]

#Add a column for the subfam (ref for the reference genes)
combined_data_pir_ht$subfam <- ifelse(rownames(combined_data_pir_ht) %in% cir_id,
                                      cir_info$subfam[match(rownames(combined_data_pir_ht), 
                                                            cir_info$Geneid)],
                                      'ref')

#Get the ref gene actual names instead of the Geneid
rownames(combined_data_pir_ht) <- ifelse(
  rownames(combined_data_pir_ht) %in% housekeeping_genes[1,],
  colnames(housekeeping_genes)[ match(rownames(combined_data_pir_ht), housekeeping_genes[1,]) ],
  rownames(combined_data_pir_ht)
)

#Get the colour range
maxcol <- ceiling(
  max(
    log2(1+combined_data_pir_ht[!colnames(combined_data_pir_ht) %in% 'subfam'])
  ))

# pir_highest_tpm_order <- rownames(combined_data_pir_ht)

#Sub-family annotation
clade_info <- combined_data_pir_ht$subfam
annotation_col <- c(clade_col, 'ref' = 'black')

#Subfamily annotation

clade_ha <- rowAnnotation(
  'sub-family' = anno_simple(
    clade_info,
    col = annotation_col,
  ),
  annotation_name_gp = gpar(fontsize = 10)
)

#Subfamily name labels
num_subfam <- cir_info %>% 
  group_by(subfam) %>% 
  summarise(num = n_distinct(Gene.ID))

labels_subfam <- c(setNames(paste0(num_subfam$subfam, ' (n = ', num_subfam$num, ')'),
                            num_subfam$subfam),
                   'ref' = 'ref')

anc_ha <- rowAnnotation(foo = anno_mark(at =  na.omit(match(clade_info, 'ancestral')) , 
                                        labels = 'ancestral',
                                        labels_gp = gpar(fontsize = 7),
                                        link_width = unit(3, "mm")))

# combined_data_pir_ht <- twentyfour_avgtpm$RMT$tpm %>% 
#   column_to_rownames(var = 'Geneid')
# 
# #Get the times instead of the stage names.
# colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
#                                               pattern = '[[:digit:]]{2}h')
# 
# combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
#                                              drop = FALSE]

ht <- Heatmap(
  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'subfam']),
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  name = 'log(TPM+1)', 
  col = colorRamp2(c(0,
                     maxcol/4,
                     maxcol/2,
                     3*maxcol/4,
                     maxcol),
                   viridis(5)),
  na_col = 'white', 
  border = TRUE,
  border_gp = gpar(lwd = 0.1),
  column_names_gp = gpar(fontsize = 10),
  row_dend_width = unit(3, "cm"),
  width = unit(4.5, 'cm'),
  height = unit(7,'cm'),
  row_title = NULL,
  left_annotation = clade_ha,
  right_annotation = anc_ha,
  row_split = forcats::fct_shift(
    relevel(
      as.factor(combined_data_pir_ht$subfam),
      ref = 'ref')
  ),
  heatmap_legend_param = list(
    labels_gp = gpar(fontsize = 10),
    title_gp = gpar(fontsize = 10,
                    fontface = 'bold'),
    at = c(0,
           maxcol/4,
           maxcol/2,
           3*maxcol/4,
           maxcol)
  )
)

clade_lgd <- Legend(
  title = 'sub-family',
  at = c(labels_subfam),
  legend_gp = gpar(fill = annotation_col)
)

draw(ht, 
     row_title_side = 'left',
     column_title_side = 'bottom',
     heatmap_legend_side = 'right',
     annotation_legend_side = 'right',
     annotation_legend_list = list(clade_lgd),
     merge_legend = TRUE)
```
